PANCHANGE_data_preprocessing

### 2020-08-13 PANCHANGE_data_preprocessing2020-07-15.Rmd
Recoded NBR ethnicity factors to match GLAD/EDGI/RAMP. Edited script to merge NBR dem and cohort datasets to use tidyverse.

# Specify cohort
```{r Specify cohort}
GLAD = FALSE
EDGI = FALSE
NBR = FALSE
RAMP = TRUE
```

#Packages
Install packages (if they are not available in your version of R)
```{r Installing packages}
#install.packages("summarytools")
#install.packages("tidyverse")
#install.packages("psych")
#install.packages("dplyr")
```

Load packages
```{r Load packages}
library(knitr)
library(summarytools)
library(psych)
library(dplyr)
library(tidyverse)
```

@@

# Select variables
```{r glad dem cleaning}
if(GLAD == TRUE){
exclude_cols_dem <- c("ID",
                  "Sample",
                  "startDate.prepandemic",
                  "Age_uncleaned",
                  "EduYrs")

@@

# Select variables
```{r edgi dem cleaning}
if(EDGI == TRUE){
  exclude_cols_dem <- c("ID",
                        "Sample",
                        "startDate.prepandemic",
                        "Age_uncleaned",
                        "EduYrs")

@@

# Select variables
```{r nbr dem cleaning}
if(NBR == TRUE){
  exclude_cols_dem <- c("ID",
                        "Sample",
                        "startDate.nbr",
                        "Age_uncleaned",
                        "EduYrs")

  nbr.dem.raw.id <- nbr.dem.raw %>%
    drop_na(subjectid) %>% # Drop NAs
    distinct(subjectid, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(subjectid, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
    separate(ID.intm, into = "ID", sep = 7) %>%
    mutate(ID = as.numeric(ID),
           startDate.coping = startDate) %>% # Rename startDate
    select(
      Sample, # Sample
      ID, # ID
      startDate.coping, #Date when participant signed up to nbr
#      Age_uncleaned = demographics.how_old_are_you_now.txt, # Age
      Birthyear = demographics.year, # Birthyear
      Gender = demographics.which_gender_do_you_identify_with, # Gender
      Sex = demographics.select_questionnaire_items_medical, #Sex
      EduYrs = demographics.school_education_school_years.txt, # Years in education
      demographics.college_or_university_degree, # level of education - univeristy
      demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
      demographics.o_levelsgcses_or_equivalent, # level of education - gcse
      demographics.cses_or_equivalent, # level of education - cse
      demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
      Ethnicity = demographics.questions_based_ethnic_origin # Ethnicity
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

```{r nbr dem cleaning}
if(NBR == TRUE) {
exclude_cols_dem <- c("ID",
                  "Sample",
                  "startDate.coping",
                  "Age_uncleaned",
                  "EduYrs")

nbr.cohort.raw.id <- nbr.cohort.raw %>%
  drop_na(`NBR ID`) %>% # Drop NAs
  distinct(`NBR ID`, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(`NBR ID`, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
  separate(ID.intm, into = "ID", sep = 7) %>%
  mutate(
    ID = as.numeric(ID),
#         startDate.coping = startDate
         ) %>% # Rename startDate
  select(
         Sample, # Sample
         ID, # ID
#         startDate.coping, #Date when participant signed up to NBR COPING
         Age_uncleaned = age, # Age
         Gender = gender, # Gender
#         Sex = demographics.select_questionnaire_items_medical, #Sex
#         EduYrs = demographics.school_education_school_years.txt, # Years in education
#         demographics.college_or_university_degree, # level of education - univeristy
#         demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
#         demographics.o_levelsgcses_or_equivalent, # level of education - gcse
#         demographics.cses_or_equivalent, # level of education - cse
#         demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         Ethnicity = ethnicity # Ethnicity
         ) %>%
  add_numeric(., exclude = exclude_cols_dem) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

```{r nbr copy dem}
if(NBR == TRUE){
dem.raw.id <- nbr.dem.raw.id
}
```

```{r nbr demo summary}
if(NBR == TRUE) {
summarytools::dfSummary(
  nbr.cohort.raw.id,
  graph.col = F,
  valid.col = F,
  labels.col = F)
}
```

# RAMP demographics
## Read in data
```{r ramp dem cols}
if(RAMP == TRUE){
  ramp.dem.raw <- readRDS(file = paste0(data.raw_path, "/ramp/dem_ramp.rds"))
  dim(ramp.dem.raw)
  colnames(ramp.dem.raw)
}
```

```{r}

@@

# Select variables
```{r ramp dem cleaning}
if(RAMP == TRUE){
  exclude_cols_dem <- c("ID",
                        "Sample",
                        "startDate.ramp")

ramp.dem.raw$externalDataReference <- ramp.dem.raw$`Login ID` #move Login ID to external data reference

  ramp.dem.raw.id <- ramp.dem.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID),
           startDate.ramp = startDate) %>% # Rename startDate
    select(
      Sample, # Sample
     # "Login ID", # ID
      startDate.ramp, #Date when participant signed up to ramp
      age_category = demographics.how_old_are_you, # Age
      Gender = demographics.which_gender_do_you_identify_with, # Gender (what [GLAD] vs which [RAMP])
     # EduYrs = demographics.university_education_attend_school.txt, # Years in education - not in RAMP dataset
     # demographics.college_or_university_degree, # level of education - univeristy - level of education is in employment file
     # demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
    #  demographics.o_levelsgcses_or_equivalent, # level of education - gcse
     # demographics.cses_or_equivalent, # level of education - cse
     # demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
      Transgender_uncleaned = demographics.do_you_identify_as_transgender,
      Ethnicity = demographics.what_is_your_ethnic_origin # Ethnicity
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

  # Inspect dimensions
  dim(ramp.dem.raw.id)
  # Inspect colnames
  colnames(ramp.dem.raw.id)
  #Differences
  dim(ramp.dem.raw)[1]-dim(ramp.dem.raw.id)[1]
}
```

```{r ramp copy dem}
if(RAMP == TRUE){
  dem.raw.id <- ramp.dem.raw.id
}
```

#Clean Age variable
Age outlier
```{r age outlier}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE ) {
# define age limits 
age_upper_limit = 117 # oldest person in the world is 117 years
age_lower_limit = 16

@@

# Inspect colnames
colnames(dem.raw.id)
}
```

@@

# Create categorical age groups per 10 years
dem.raw.id <- dem.raw.id %>%
  mutate(
    age_category_numeric =
      case_when(
          Age >= 16 & Age <= 18 ~ "1",
          Age >= 19 & Age <= 25 ~ "2",
          Age >= 26 & Age <= 35 ~ "3",
          Age >= 36 & Age <= 45 ~ "4",
          Age >= 46 & Age <= 55 ~ "5",
          Age >= 56 & Age <= 65 ~ "6",
          Age >= 66 & Age <= 70 ~ "7",
          Age >= 71 & Age <= 75 ~ "8",
          Age >= 76 & Age <= 80 ~ "9",
          Age >= 81 & Age <= 85 ~ "10",
          Age >= 86 & Age <= 90 ~ "11",
          Age >= 91 & Age <= 100 ~ "12",
          Age >= 101 & Age <= 120 ~ "13" # oldest person in the world is 117 years
                )
      )
}

dem.raw.id <- dem.raw.id %>%
    mutate(
      age_category_full =
        recode_factor(
          age_category_numeric, 
          "1" = "16 to 18 years",
          "2" = "19 to 25 years",
          "3" = "26 to 35 years",
          "4" = "36 to 45 years",
          "5" = "46 to 55 years",
          "6" = "56 to 65 years",
          "7" = "66 to 70 years",
          "8" = "71 to 75 years",
          "9" = "76 to 80 years",
          "10" = "81 to 85 years",
          "11" = "86 to 90 years",
          "12" = "91 to 100 years",
          "13" = "Above 100 years"
          )
    )

dem.raw.id %>%
  freq(age_category_full)

# RAMP employment
## Read in data
```{r ramp emp cols}
if(RAMP == TRUE){
  ramp.emp.raw <- readRDS(file = paste0(data.raw_path, "/ramp/employment_ramp.rds"))
  dim(ramp.emp.raw)
  colnames(ramp.emp.raw)
}
```

# Select variables
```{r ramp emp cleaning}
if(RAMP == TRUE){
  exclude_cols_emp <- c("ID",
                        "Sample")

ramp.emp.raw$externalDataReference <- ramp.emp.raw$`Login ID` #move Login ID to external data reference

ramp.emp.raw.id <- ramp.emp.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID),
           startDate.ramp = startDate) %>% # Rename startDate
    select(
      highest_education_ramp = employment.what_is_your_highest_level_of_education,
      # EduYrs = employment.university_education_attend_school.txt, # Years in education - not in RAMP dataset
      employment.government_work_key_workers
    ) %>%
    add_numeric(., exclude = exclude_cols_emp) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

  # Inspect dimensions
  dim(ramp.emp.raw.id)
  # Inspect colnames
  colnames(ramp.emp.raw.id)
  #Differences
  dim(ramp.emp.raw)[1]-dim(ramp.emp.raw.id)[1]
}
```

```{r ramp copy emp}
if(RAMP == TRUE){
  emp.raw.id <- ramp.emp.raw.id
}
```

# Highest education
Proceed along qualifications fields. If field is not blank, take that as highest qual. If field is blank, check next field. If all fields are blank, NA
Assumes University degree > A level > NVQ > GCSE/O Level/CSE
```{r Highest education}
if(GLAD == TRUE | EDGI == TRUE) {
#create numeric version of the highest education variable
dem.raw.id <- dem.raw.id %>%
  mutate(
    highest_education_numeric =
      case_when(
        demographics.college_or_university_degree_numeric == "1" ~ 4,
        demographics.a_levelsas_levels_or_equivalent_numeric == "1" ~ 3,
        demographics.nvq_or_hnd_or_hnc_or_equivalent_numeric == "1" ~ 2,
        demographics.o_levelsgcses_or_equivalent_numeric == "1" ~ 1,
        demographics.cses_or_equivalent_numeric == "1" ~ 1)
    )

@@

Check highest education coding
```{r Highest education check}
if(GLAD == TRUE | EDGI == TRUE) {
dem.raw.id %>%
  select(highest_education,
        demographics.college_or_university_degree_numeric,
        demographics.a_levelsas_levels_or_equivalent_numeric,
        demographics.nvq_or_hnd_or_hnc_or_equivalent_numeric,
        demographics.o_levelsgcses_or_equivalent_numeric,
        demographics.cses_or_equivalent_numeric
  )
}
```

@@

# GLAD employment
## Read in data
```{r glad dem cols}
if(GLAD == TRUE){
glad.employment.raw <- readRDS(file = paste0(data.raw_path, "/glad/employment_coping_glad.rds"))
dim(glad.employment.raw)
colnames(glad.employment.raw)
}
```

@@

>>>>

# Specify cohort
```{r Specify cohort}
GLAD = FALSE
EDGI = FALSE
NBR = TRUE
RAMP = FALSE
```

#Packages
Install packages (if they are not available in your version of R)
```{r Installing packages}
#install.packages("summarytools")
#install.packages("tidyverse")
#install.packages("psych")
#install.packages("dplyr")
#install.packages("eeptools")
```

Load packages
```{r Load packages}
library(knitr)
library(summarytools)
library(psych)
library(dplyr)
library(tidyverse)
library(eeptools)
```

@@

## Select variables
```{r glad dem cleaning}
if(GLAD == TRUE){
exclude_cols_dem <- c("ID",
                  "Sample",
                  "startDate.prepandemic",
                  "Age_uncleaned",
                  "EduYrs")

@@

## Select variables
```{r edgi dem cleaning}
if(EDGI == TRUE){
  exclude_cols_dem <- c("ID",
                        "Sample",
                        "startDate.prepandemic",
                        "Age_uncleaned",
                        "EduYrs")

@@

## Select variables
```{r nbr dem cleaning}
if(NBR == TRUE){
  exclude_cols_dem <- c("ID",
                        "Sample",
                        "startDate.nbr",
                        "Age_uncleaned",
                        "EduYrs")

  nbr.dem.raw.id <- nbr.dem.raw %>%
    drop_na(subjectid) %>% # Drop NAs
    distinct(subjectid, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(subjectid, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
    separate(ID.intm, into = "ID", sep = 7) %>%
    mutate(ID = as.numeric(ID),
           startDate.coping = startDate) %>% # Rename startDate
    select(
      Sample, # Sample
      ID, # ID
      startDate.coping, #Date when participant signed up to nbr
#      Age_uncleaned = demographics.how_old_are_you_now.txt, # Age
      Birthyear_unc = demographics.year, # Birthyear
      Birthmonth = demographics.month, #Birthmonth
      Birthday = demographics.day, #Birthday
      Gender = demographics.which_gender_do_you_identify_with, # Gender
      Sex = demographics.select_questionnaire_items_medical, #Sex
#      Ethnicity = demographics.questions_based_ethnic_origin # Ethnicity
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

```{r nbr cohort cleaning}
if(NBR == TRUE) {
exclude_cols_dem <- c("ID",
                  "Sample",
                  "startDate.coping",
                  "Age_uncleaned",
                  "EduYrs")

nbr.cohort.raw.id <- nbr.cohort.raw %>%
  drop_na(`NBR ID`) %>% # Drop NAs
  distinct(`NBR ID`, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(`NBR ID`, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
  separate(ID.intm, into = "ID", sep = 7) %>%
  mutate(
    ID = as.numeric(ID),
#         startDate.coping = startDate
         ) %>% # Rename startDate
  select(
         Sample, # Sample
         ID, # ID
         Cohort = cohort, #cohort information
#         startDate.coping, #Date when participant signed up to NBR COPING
#         Age_uncleaned = age, # Age
#         Gender = gender, # Gender
#         Sex = demographics.select_questionnaire_items_medical, #Sex
#         EduYrs = demographics.school_education_school_years.txt, # Years in education
#         demographics.college_or_university_degree, # level of education - univeristy
#         demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
#         demographics.o_levelsgcses_or_equivalent, # level of education - gcse
#         demographics.cses_or_equivalent, # level of education - cse
#         demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         Ethnicity.unc = ethnicity # Ethnicity
         ) %>%
  add_numeric(., exclude = exclude_cols_dem) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

```{r nbr cohort recode ethnicity}
#Recode details ethnicity info
nbr.cohort.raw.id$Ethnicity_full <- recode_factor(nbr.cohort.raw.id$Ethnicity.unc,
                                             "A" = "White - British",
                                             "B" = "White - Irish",
                                             "C" = "White - Other",
                                             "D" = "Mixed - White & Black Caribbean",
                                             "E" = "Mixed - White & Black African",
                                             "F" = "Mixed - White & Asian",
                                             "G" = "Mixed - Other",
                                             "H" = "Asian - Indian",
                                             "J" = "Asian - Pakistani",
                                             "K" = "Asian - Bangladeshi",
                                             "L" = "Asian - Other",
                                             "M" = "Black - Caribbean",
                                             "N" = "Black - African",
                                             "P" = "Black - Other",
                                             "R" = "Chinese - Chinese",
                                             "S" = "Chinese - Other",
                                             "W" = NULL, ## Asking NBR what this level means
                                             "Z" = NULL,
                                             "99" = NULL)

#Match GLAD/EDGI/RAMP categories
nbr.cohort.raw.id$Ethnicity <- recode_factor(nbr.cohort.raw.id$Ethnicity.unc,
                                             "A" = "White, white European or Caucasian",
                                             "B" = "White, white European or Caucasian",
                                             "C" = "White, white European or Caucasian",
                                             "D" = "Mixed or multiple ethnic origins",
                                             "E" = "Mixed or multiple ethnic origins",
                                             "F" = "Mixed or multiple ethnic origins",
                                             "G" = "Mixed or multiple ethnic origins",
                                             "H" = "Asian or Asian British (including Chinese)",
                                             "J" = "Asian or Asian British (including Chinese)",
                                             "K" = "Asian or Asian British (including Chinese)",
                                             "L" = "Asian or Asian British (including Chinese)",
                                             "M" = "Black or Black British",
                                             "N" = "Black or Black British",
                                             "P" = "Black or Black British",
                                             "R" = "Asian or Asian British (including Chinese)",
                                             "S" = "Asian or Asian British (including Chinese)",
                                             "W" = NULL, ## Asking NBR what this level means
                                             "Z" = NULL,
                                             "99" = NULL)

@@

```{r nbr combine demographics}
if(NBR == TRUE){
dem.raw.id <- left_join(nbr.dem.raw.id, nbr.cohort.raw.id, by = c("Sample", "ID"))

# Inspect dimensions
dim(dem.raw.id)
# Inspect colnames
colnames(dem.raw.id)
#Differences
dim(nbr.dem.raw.id)[1]-dim(dem.raw.id)[1]
}
```

```{r nbr demo summary}
if(NBR == TRUE) {
summarytools::dfSummary(
  dem.raw.id,
  graph.col = F,
  valid.col = F,
  labels.col = F)
}
```

#Clean Age variable
Age outlier
```{r age outlier}
# define age limits 
age_upper_limit = 117 # oldest person in the world is 117 years
age_lower_limit = 16

@@

# Inspect colnames
colnames(dem.raw.id)
```

@@

# Create categorical age groups per 10 years
dem.raw.id <- dem.raw.id %>%
  mutate(
    age_category_numeric =
      case_when(
          Age >= 16 & Age <= 18 ~ "1",
          Age >= 19 & Age <= 25 ~ "2",
          Age >= 26 & Age <= 35 ~ "3",
          Age >= 36 & Age <= 45 ~ "4",
          Age >= 46 & Age <= 55 ~ "5",
          Age >= 56 & Age <= 65 ~ "6",
          Age >= 66 & Age <= 70 ~ "7",
          Age >= 71 & Age <= 75 ~ "8",
          Age >= 76 & Age <= 80 ~ "9",
          Age >= 81 & Age <= 85 ~ "10",
          Age >= 86 & Age <= 90 ~ "11",
          Age >= 91 & Age <= 100 ~ "12",
          Age >= 101 & Age <= 120 ~ "13" # oldest person in the world is 117 years
                )
      )

dem.raw.id <- dem.raw.id %>%
    mutate(
      age_category =
        recode_factor(
          age_category_numeric, 
          "1" = "16 to 18 years",
          "2" = "19 to 25 years",
          "3" = "26 to 35 years",
          "4" = "36 to 45 years",
          "5" = "46 to 55 years",
          "6" = "56 to 65 years",
          "7" = "66 to 70 years",
          "8" = "71 to 75 years",
          "9" = "76 to 80 years",
          "10" = "81 to 85 years",
          "11" = "86 to 90 years",
          "12" = "91 to 100 years",
          "13" = "Above 100 years"
          )
    )

dem.raw.id %>%
  freq(age_category)

}
```

# Highest education (prepandemic)
Proceed along qualifications fields. If field is not blank, take that as highest qual. If field is blank, check next field. If all fields are blank, NA
Assumes University degree > A level > NVQ > GCSE/O Level/CSE
```{r Highest education prepandemic}
if(GLAD == TRUE | EDGI == TRUE) {
#create numeric version of the highest education variable
dem.raw.id <- dem.raw.id %>%
  mutate(
    highest_education_numeric =
      case_when(
        demographics.college_or_university_degree_numeric == "1" ~ 4,
        demographics.a_levelsas_levels_or_equivalent_numeric == "1" ~ 3,
        demographics.nvq_or_hnd_or_hnc_or_equivalent_numeric == "1" ~ 2,
        demographics.o_levelsgcses_or_equivalent_numeric == "1" ~ 1,
        demographics.cses_or_equivalent_numeric == "1" ~ 1)
    )

@@

Check highest education coding
```{r Highest education prepandemic check}
if(GLAD == TRUE | EDGI == TRUE) {
dem.raw.id %>%
  select(highest_education,
        demographics.college_or_university_degree_numeric,
        demographics.a_levelsas_levels_or_equivalent_numeric,
        demographics.nvq_or_hnd_or_hnc_or_equivalent_numeric,
        demographics.o_levelsgcses_or_equivalent_numeric,
        demographics.cses_or_equivalent_numeric
  )
}
```

@@

# Employment/highest education (baseline)

## Read in data
```{r glad employment cols}
if(GLAD == TRUE){
glad.employment.raw <- readRDS(file = paste0(data.raw_path, "/glad/employment_coping_glad.rds"))
dim(glad.employment.raw)
colnames(glad.employment.raw)
}
```

@@

```{r nbr employment cols}
if(NBR == TRUE){
nbr.employment.raw <- readRDS(file = paste0(data.raw_path, "/nbr/employment_coping_nbr.rds"))
dim(nbr.employment.raw)
colnames(nbr.employment.raw)
}
```

# Select variables
```{r nbr employment cleaning}
if(NBR == TRUE){
exclude_cols_employment <- c("ID",
                  "Sample")

nbr.employment.raw.id <- nbr.employment.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(
    ID = as.numeric(ID)
    ) %>%
  select(
         Sample, # Sample
         ID, # ID
         employment.phd, # level of education - univeristy
         employment.masters_degree_or_equivalent, # level of education - univeristy
         employment.postgraduate_degree_or_equivalent, # level of education - univeristy
         employment.bachelors_degree_or_equivalent, # level of education - univeristy
         employment.a_levelsas_levels_or_equivalent, # level of education - alevel
         employment.o_levelsgcses_or_equivalent, # level of education - gcse
         employment.cses_or_equivalent, # level of education - cse
         employment.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         employment.government_work_key_workers # key worker status
         ) %>%
  add_numeric(., exclude = exclude_cols_employment) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(nbr.employment.raw.id)
# Inspect colnames
colnames(nbr.employment.raw.id)
#Differences
dim(nbr.employment.raw)[1]-dim(nbr.employment.raw.id)[1]
}
```
